---
title: "inclass"
author: "Rithin"
date: "3/3/2020"
output: html_document
---

```{r}
library(tibble)
library(readr)

data_dir <- "~/Downloads/jena_climate"
fname <- file.path(data_dir, "jena_climate_2009_2016.csv")
data <- read_csv(fname)
```

## data overview

```{r}
glimpse(data)
```

Plot of temperatur over time. In this plot we can observe the yearly periodicity of temperature.

```{r}
library(ggplot2)
ggplot(data, aes(x = 1:nrow(data), y = `T (degC)`)) + geom_line()
```

## Preparing the data ##

```{r}
data <- data.matrix(data[,-1])
```

## Normalizing the data ##

```{r}
train_data <- data[1:200000,]
mean <- apply(train_data, 2, mean)
std <- apply(train_data, 2, sd)
data <- scale(data, center = mean, scale = std)
```

```{r}
generator <- function(data, lookback, delay, min_index, max_index,
                      shuffle = FALSE, batch_size = 128, step = 6) {
  if (is.null(max_index))
    max_index <- nrow(data) - delay - 1
  i <- min_index + lookback
  function() {
    if (shuffle) {
      rows <- sample(c((min_index+lookback):max_index), size = batch_size)
    } else {
      if (i + batch_size >= max_index)
        i <<- min_index + lookback
      rows <- c(i:min(i+batch_size-1, max_index))
      i <<- i + length(rows)
    }
    
    samples <- array(0, dim = c(length(rows), 
                                lookback / step,
                                dim(data)[[-1]]))
    targets <- array(0, dim = c(length(rows)))
                     
    for (j in 1:length(rows)) {
      indices <- seq(rows[[j]] - lookback, rows[[j]] - 1, 
                     length.out = dim(samples)[[2]])
      samples[j,,] <- data[indices,]
      targets[[j]] <- data[rows[[j]] + delay,2]
    }            
    
    list(samples, targets)
  }
}
```

```{r}
lookback <- 1440
step <- 6
delay <- 144
batch_size <- 128

train_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 1,
  max_index = 200000,
  shuffle = TRUE,
  step = step, 
  batch_size = batch_size
)

val_gen = generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 200001,
  max_index = 300000,
  step = step,
  batch_size = batch_size
)

test_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 300001,
  max_index = NULL,
  step = step,
  batch_size = batch_size
)

## to look at thw whole validation data set:
val_steps <- (300000 - 200001 - lookback) / batch_size

# in order to see the whole test set:
test_steps <- (nrow(data) - 300001 - lookback) / batch_size
```

```{r}
evaluate_naive_method <- function() {
  batch_maes <- c()
  for (step in 1:val_steps) {
    c(samples, targets) %<-% val_gen()
    preds <- samples[,dim(samples)[[2]],2]
    mae <- mean(abs(preds - targets))
    batch_maes <- c(batch_maes, mae)
  }
  print(mean(batch_maes))
}
```

## Model - 1 ##

I used gru layer on the top with 32 units
second layer was dense with number of units1

```{r}
model <- keras_model_sequential() %>% 
  layer_gru(units = 32, input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```

```{r}
plot(history)
```

The above model was overfitting, So an approach to overcome this problem was droup out.

## Model - 2 ##

Using gru with 32 units with dropout of 0.2.
We used drop out to reduce overfitting.
using dense layer with 1 unit.

```{r}
model <- keras_model_sequential() %>% 
  layer_gru(units = 32, dropout = 0.2, recurrent_dropout = 0.2,
            input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 40,
  validation_data = val_gen,
  validation_steps = val_steps
)
```

```{r}
plot(model)
```

we are no longer overfitting during the first 30 epochs. 

## Model - 3 ##

Using LSTM insead of gru with 16 units in first layer and 32 units in second layer with drop out of 0.1
using dense layer of units 1

```{r}
model1 <- keras_model_sequential() %>% 
  layer_lstm(units = 16, 
            dropout = 0.1, 
            recurrent_dropout = 0.5,
            return_sequences = TRUE,
            input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_lstm(units = 32, activation = "relu",
            dropout = 0.1,
            recurrent_dropout = 0.5) %>% 
  layer_dense(units = 1)

model1 %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history1 <- model1 %>% fit_generator(
  train_gen,
  steps_per_epoch = 100,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```

```{r}
plot(history1)
```

## Model - 4 ##

Using LSTM layer with 32 and 64 units in first and second layer with drop out of 0.1
Also using dense layer of unit 1

```{r}

model1 <- keras_model_sequential() %>% 
  layer_lstm(units = 32, 
            dropout = 0.1, 
            recurrent_dropout = 0.5,
            return_sequences = TRUE,
            input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_lstm(units = 64, activation = "relu",
            dropout = 0.1,
            recurrent_dropout = 0.5) %>% 
  layer_dense(units = 1)

model1 %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history1 <- model1 %>% fit_generator(
  train_gen,
  steps_per_epoch = 100,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)

```

```{r}
plot(history1)
```

This model is best because we have stacked recurrent layers and reduced overfitting by adding droupouts.
